---
title: "Tidy and clean Linkedin Scraped Data"
author: Anil Akyildirim, Nicholas Chung, Jai Jeffryes, Tamiko Jenkins, Joe Rovalino,
  Sie Siong Wong
date: "10/20/2019"
output:
  html_document:
    code_download: yes
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

## Introduction

As part of our project, we are tasked to answer the question "What are the most valued data science skills?" by working as a team, deciding what data to collect and how to collect it, use relational database and set of normalized tables and data exploration and analysis. Our team members are as follows;

- Anil Akyildirim

- Nicholas Chung

- Jai Jeffryes

- Tamiko Jenkins

- Joe Rovalino

- Sie Siong Wong

As part of project management tools, we have used Slack Private channel and Skype for Project Communication, Github for Project tracking, documentation and code collaboration, and Amazon Relational Database Service for data integration. All of our supporting code and data are on the GitHub repo, which documents branches and commits from our team.

- GitHub: [https://github.com/pnojai/dskill](https://github.com/pnojai/dskill)
- Amazon Relational Database Service: [msds607.ckxhi71v1dqf.us-east-1.rds.amazonaws.com](msds607.ckxhi71v1dqf.us-east-1.rds.amazonaws.com)


## Load Data {.tabset}

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# load packages
library(dplyr)
library(jsonlite)
library(ggplot2)
library(RMariaDB)
```

### Data Collection 

We have reviewed and discussed different data types such as current job requirements around data scientists from job postings such as indeed.com or monster.com and articles around top data scientists skills in websites such as towardsdatascience and knuggets. Our approach built on the assumption that data scientists with jobs have the skills most valued by employers. We collected skills from employed data scientists.

We were inspired by the research of Jeff Hale whose article on data science skills appeared on the website, Medium.

- [https://towardsdatascience.com/the-most-in-demand-skills-for-data-scientists-4a4a8db896db](https://towardsdatascience.com/the-most-in-demand-skills-for-data-scientists-4a4a8db896db).

We discussed different methods of collecting the data and further how we can store it. As a result, we decided to work with usefull data within linkedin.com. We compared our findings from LinkedIn data to Mr. Hale's 2018 findings.


### Load JSON files
```{r}
# load all JSON
filenames <- list.files("data/profiles", pattern="*.json", full.names=TRUE) # this should give you a character vector, with each file name represented by an entry
example_file <- lapply(filenames[1], function(x) jsonlite::fromJSON(txt = x)) # a list in which each element is one of your original JSON files
example_file
```

## Structure Data {.tabset}

### Bind fromJSON results
```{r}
# apply fromJSON to read in all of the json files
# create the column (variable) title, headline, which will be populated with json file identifying information
# extract the skills data which contains the variables title and counts
# bind the results together as a data frame named raw_df
raw_df <- dplyr::bind_rows(sapply(filenames, function(x) fromJSON(x, flatten=TRUE)$skills), .id="headline")
head(raw_df)
```


### Extract Headlines
```{r}
# apply fromJSON to read in all of the json files
# extract the headline variable from the profile data, saving each file name as the variable title
# save the mapping as  data frame headlines 
headlines <- sapply(filenames, function(x) fromJSON(x, flatten=TRUE)$profile$headline, USE.NAMES = TRUE)
head(headlines)
```


### Map Headlines to Skills
```{r}
#  apply a look up of the variable title specifying the filename 
#  and add the headline value from the headlines  data frame
#  to the headlines variable in  data frame raw_df

raw_df$headline <- sapply(raw_df$headline, function(x) headlines[x])
head(raw_df)
```

## Format Data {.tabset}

### Name and convert variables and data
```{r}
df_conv <- raw_df
names(df_conv) <- c("title", "skills", "count")
class(df_conv)
# coerce any nulls to na's
sapply(df_conv, class)
# create numeric types in counts column
df_conv$count <- as.numeric(df_conv$count)
sapply(df_conv, class)
```


### Remove NA's from numeric column
```{r}
# count all rows
# 4822
nrow(df_conv)
# view a subset of rows with na's mixed with complete rows
df_conv[11:20,]

# filter for any rows with na
# count all rows with na's
# 942
df_na <- df_conv %>% filter_all(any_vars(is.na(.)))
nrow(df_na)

# omit any rows with na's
# save rows without na's as a data frame names df
# count the data frame 
# 4822-942 = 3880
df_omit <- na.omit(df_conv)
# view the same rows without na's
df_omit[11:20,]

nrow(df_omit)
```


## Store Data {.tabset}

### Prepare data values for storage
```{r}
#  Encoding(x) <- "latin1"
#  x <-  iconv(x, "latin1", "UTF-8", sub='')
#  x <- stringr::str_replace(x,",","")
#  Encoding(x) <- "UTF-8"


df_clean <- df_omit
# Remove non-ASCII character codes
test <- df_clean[256,]
df_clean$skills <- sapply(df_clean$skills, function(x) gsub('[^\x20-\x7E]', '', x))
df_clean$title <- sapply(df_clean$title, function(x) gsub('[^\x20-\x7E]', '', x))
df_clean$skills <- sapply(df_clean$skills, function(x) gsub('[@]', 'at', x))
df_clean$title <- sapply(df_clean$title, function(x) gsub('[@]', 'at', x))
df_clean$skills <- sapply(df_clean$skills, function(x) gsub('[\\|\\(\\),]', '', x))
df_clean$title <- sapply(df_clean$title, function(x) gsub('[\\|\\(\\),]', '', x))
Encoding(df_clean$skills) <- "UTF-8"
Encoding(df_clean$title) <- "UTF-8"
head(df_clean)

# View original
test
# view cleaned text
df_clean[256,]
```


### Prepare data format for storage
```{r}

# TODO: follow df.csv convention

# Add rownames (indices) as a skill id
# to final dataframe to prepare for 
# SQL-based storage and to provide option to
# remove automatic row names from write csv
# Remove depr function
#df_csv <- add_rownames(df, var = "skill_id")

# NB: these are the original row id's based on R records
# to generate skill_ids without skips for na's removed
# use a seq
df_csv <- tibble::rownames_to_column(df_clean, var = "skill_id")
df_csv$skill_id <- as.numeric(df_csv$skill_id)
head(df_csv)

# TODO: follow df.csv convention
# Rearrange column order with dplyr select
df_csv <- dplyr::select(df_csv, skill_id, skills, count, title)

head(df_csv)

```



### Write to csv
```{r}
# write csv and upload to our mysql database
# Encoding(df_csv)
write.csv(df_csv, "results/df_alt.csv", row.names=FALSE, fileEncoding="UTF-8")

```



## Load Data {.tabset}


### Load the data from the database
```{r}
# load the data in the database and look at 2018 Linkedin Data
user_name <- 'anil'
user_password <- "redy2rok"
database <- 'prj3'
host_name <- 'msds607.ckxhi71v1dqf.us-east-1.rds.amazonaws.com'

#connecting to the MySQL database

myDb <- dbConnect(RMariaDB::MariaDB(), user=user_name, password=user_password, dbname=database, host=host_name)
myDb
```

### View tables

```{r}
#list of tables we have
dbListTables(myDb)

```




## Exploring the Data  {.tabset}

### View 2018 Data
```{r}
# lets load 2018 Linkedin Data
df <- dbGetQuery(myDb, "select * from df")
head(df)
nrow(df)

```

### View 2019 Data

```{r}
# There are more than 145 skills, clean to data similar to 2018 data
df <- subset(df, select = c(skills, count))
colnames(df) <- c("Skills", "Linkedin")
head(df)
nrow(df)

```

```{r}
# there are skills that is listed more than once. finding those
n_occur <- data.frame(table(df$Skills))
skills_more_once <- n_occur[n_occur$Freq > 1,]
head(skills_more_once)
nrow(skills_more_once)
```

```{r}
# we need to add the count of the duplicate skills rows 

df <- aggregate(Linkedin ~ Skills, dat=df, FUN=sum)
head(df)

```

```{r}
# data is collected and ready to be analyzed at this point
summary(df)
str(df)

```

```{r}

df$Skills <- as.character(df$Skills)
df$Linkedin <- as.numeric(df$Linkedin)

```



## Visualizing the Data 

### Improvements

The first exploratory pass is crowded. We'll filter the data in the next pass.

```{r}

#We have 1157 observations (skills) that data science roles use in linkedin
#let's see the distribution

theme_set(theme_classic())

ggplot(df, aes(x=Skills, y=Linkedin))+
  geom_bar(stat="identity", width = 0.5, fill=("tomato2"))+
             theme(axis.text.x = element_text(angle = 65, vjust=0.6))


```

```{r}
# we have way too many skills so let's only focus on the ones that has significant count.

df <- filter(df, Linkedin >100)
head(df)
nrow(df)

```

### Visualization
```{r}
# we narrowed it down to 57 skills. Let's see how distribution looks like.

theme_set(theme_classic())

ggplot(df, aes(x=Skills, y=Linkedin))+
  geom_bar(stat="identity", width = 0.5, fill=("tomato2"))+
             theme(axis.text.x = element_text(angle = 65, vjust=0.6))



```

```{r}

# let's narrow it down further. 
df <- filter(df, Linkedin > 200)
head(df)
nrow(df)
```


```{r}

theme_set(theme_classic())

ggplot(df, aes(x=reorder(Skills, Linkedin, fun=max), y=Linkedin))+
  geom_bar(stat="identity", width = 0.5, fill=("tomato2"))+
  labs(title="2019 Data Science Skills Distribution",
       x="Data Science Skills",
       y="Count in Linkedin")+
  theme(axis.text.x = element_text(angle = 65, vjust=0.6))


```


### Analyze the Data 

The Data Science skills Distribution chart for 2019 shows us the most frequent data science skills that people use for their Linkedin Profiles. The results show us that, Data Analysis; as part of General Data Skills, is the most commonly used skill within Data Scientists in Linkedin. The top three programming languages used within the profiles are R, Python and SQL. Statistics and Machine Learning are in 5th and 6th place in that order. If we consider Machine Learning and Statistics, as part of General Data Science Skills and Programming Languages as part of Technical Data Science Skills, we can conclude that 

**Top three General Data Science Skills are Data Analysis, Statistics and Machine Learning.**

**Top three Technical Data Science Skills are R, Python and SQL.**

```{r}
# count for top three General and Technical Data Science Skills

data_analysis <- filter(df, df$Skills=="Data Analysis")
machine_learning <- filter(df, df$Skills=="Machine Learning")
statistics <- filter(df, df$Skills=="Statistics")
python <- filter(df, df$Skills=="Python")
r <- filter(df, df$Skills=="R")
sql <- filter(df, df$Skills=="SQL")
data_analysis
machine_learning
statistics
python
r
sql

```



We can also look at the LinkedIn Data Set from Jeff Hale and see if they follow the same pattern.

```{r}
# lets load 2018 Linkedin Data from Jeff Hale.
skills_2018 <- dbGetQuery(myDb, "select * from ds_general_skills_clean")
skills_2018$LinkedIn <- as.numeric(skills_2018$LinkedIn) # little cleanup
skills_2018


```


```{r}
# analyze briefly to see if there are differences
theme_set(theme_classic())

ggplot(skills_2018, aes(x=reorder(Keyword, LinkedIn, fun=max),y=LinkedIn))+
  geom_bar(stat="identity", width = 0.5, fill=("tomato2"))+
  labs(title="2018 Data Science Skills Distribution",
       x="Data Science Skills",
       y="Count in Linkedin",
       caption = "Source: Jeff Hale 2018 Data Skills Analysis")+
  theme(axis.text.x = element_text(angle = 65, vjust=0.6))

```

With the assumption of computer science covering the Programming Lanaguages, we can see that the data science skills distribution for 2018 is similar to our Data Science Skills Distribution for 2018. Machine Learning , Data Analysis and Statistics leading the top General Data Science Skills. The only difference we see is that Machine Learning is slightly more used Data Science Skill than Data Analysis.

### Simplify 2019 data frame

```{r}
# subset vector v, removing the headline variable[c(-1)]
# save the new vector as v_counts
df_counts <- df_csv
head(df_counts)

```

### View aggregate 2019 Skills counts
```{r}

# aggregate the counts for each unique skill
# store as agg_df_counts data frame
agg_df_counts <- df_counts %>%
    group_by(skills)  %>%
    dplyr::summarise(count = n())  %>%
    arrange(desc(count))

agg_df_counts
```


## Conclusions

The six data science skills most valued by employers in 2019 appear to be the following.

General Data Science Skills:

1- Data Analysis => 2196

2- Machine Learning => 979

3- Statistics => 1036

Technical Data Science Skills

1- R => 1864

2- Python => 1539

3- SQL => 1237

Our approach differed from Mr. Hale's. He investigated programming languages as a separate research question. Our approach commingles them. Therefore, though our high-ranking skills list includes the languages R, Python, and SQL, nothing is to be concluded from their absence from Hale's list. What we see in common are the skills of analysis, statistics, and machine learning. We believe the data tell a compelling story about investment in these disciplines.



